---
title: "**The HEAT Study in its final year**" 
subtitle: "Learning more about building science than I would've ever thought😅"
author: Joe Simeone ([Theme by Grant McDermott](https://github.com/grantmcdermott/quarto-revealjs-clean))
format: 
    revealjs:
        theme: clean.scss
        slideNumber: true
---


```{r}
#| include: false

library(duckdb)
library(tidyverse)
library(ggtext)
library(tinytable)
library(glue)
library(reactable)
library(htmltools)
library(arrow)
library(scales)

# con <- dbConnect(duckdb(), 'C:/git/nrel_stuff/data/nrel.duckdb', read_only = TRUE)
# nrel_new_con <- dbConnect(duckdb(), 'C:/git/nrel_stuff/granular_workflows/data/nrel_upg_32.duckdb', read_only = TRUE)


# dbListTables(con)
# dbGetQuery(con, "SHOW TABLES") # shows tables in current schema
# dbGetQuery(con, "SELECT * FROM information_schema.tables")


#source('helper_scripts/make_indoor_temp_graphs.R')
source('helper_scripts/make_building_char_table.R')
source('helper_scripts/01_hi_plot_helpers.R')
source('helper_scripts/make_mixed_humid_hi_plot.R')
source('helper_scripts/make_philly_hi_plot.R')


tsp_theme <-
  theme_gray(base_size = 16) +   # <-- add this as a base
theme(
    panel.background = element_rect(fill = "transparent", colour = NA),
    plot.background = element_rect(fill = "transparent", colour = NA),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.background = element_rect(fill = "transparent"),
    legend.key = element_rect(fill = "transparent"),
    legend.position = 'none',
     strip.background = element_rect(fill = "transparent", color = NA),
     strip.text = element_markdown(),
     plot.title = element_markdown(size = 12, hjust = .5),
     plot.subtitle = element_markdown(size = 12, hjust = .5),
     plot.caption = element_text(size = 12)
  )

```

## Outline

::: {.incremental}
1. Brief HEAT Study refresher
2. NLR ResStock data source
3. Goals for the year  
:::


## Heat Study Refresher {.smaller}

:::: {.columns}
::: {.column width="50%"}
::: {.incremental}
- The **HEAT Study** is an NIH R01 awarded to PI Leah Schniasi to study associations between outdoor and indoor temperature and mortality in the United States.
- **Aim 1:** *Quantify and compare associations of seasonal **indoor** and outdoor temperatures with **all cause**, **cardiovascular**, **respiratory**
and **renal disease** mortality (nationally)*.

:::
:::

::: {.column width="50%"}
::: {.fragment}
### Rationale

::: {.incremental}
1. We have pretty good evidence that more people die when it's really hot or really cold outside.
1. These deaths are not (just) due to exposure, but the exacerbation of underlying chronic conditions.
1. While the associations are well-documented, the causal pathways linking extreme outdoor temperature to chronic disease mortality remain unclear.
:::

:::
:::
::::

::: {.fragment}
One thing that likely affects this relationship is the **indoor environment and housing quality.** Temperature controls (or lack thereof) will influence the actual temperature exposure a person experiences. This is especially true because the population most vulnerable to hot and cold (people aged 65 or over) spends most of their time indoors.
:::

## 


:::: {.columns}
::: {.column width="50%"}
#### But we quickly ran into a (pretty big) problem...
::: {.fragment}
> No good observational data on indoor temperature and housing exist at the longitudinal and geographic scales needed to achieve our study aims.
:::
:::

::: {.column width="50%"}
::: {.incremental}
- Basic housing information is available from the Census **Public Use Microdata Sample (PUMS)** of the American Community Survey
- Much more detailed information from the **Residential Energy Consumption Survey (RECS)**, but RECS is only representative at coarse geographies (census regions or states depending on the survey year).
:::
:::
::::

## Introducing the National Laboratory of the Rockies (NLR) ResStock Model

::: {.incremental}
- ResStock is [described](https://natlabrockies.github.io/ResStock.github.io/) as a "highly granular, bottom-up model that uses multiple data sources, statistical sampling methods, and advanced building energy simulations" to provide data on energy use in residential building stock across the United States."
- In practice, the model produces 550,000 synthetic housing simulations (each representing roughly 254 homes) drawn from conditional probability distributions to represent residential housing stock as of 2019.
- For each synthetic unit, energy use endpoints and the indoor environment are simulated at 15-minute time scales for a single simulation year (2018).
::: 

## What goes into a building simulation? {.smaller}

```{r}
#| echo: false
#| fig-cap: '*NLR ResStock 2025.1: Delaware & Philadelphia prototypes used in temperature simulations. Simulation inputs come from 3 broad categories: 1) **Weather stations** assigned to county; 2) **Physical building characteristics**; 3) **Occupant information and behaviors***'

bldg_char_table
```

## How do they overcome the key data limitation? 



<style>
.small-caption figcaption { font-size: 16px; line-height: 1.2; }
</style>

::: {.small-caption}
![*Adapted from the ResStock [Technical Documentation](https://oedi-data-lake.s3.amazonaws.com/nrel-pds-building-stock/end-use-load-profiles-for-us-building-stock/2025/resstock_amy2018_release_1/ResStockTechnicalReferenceGuide_2025_1.pdf). Housing characteristics are assigned by "dependent sampling chains": each is drawn from nested conditional probability distributions built from national survey data (ACS PUMS, RECS, AHS, ATUS) plus a few more granular sources. Users supply a quota, and housing units are sampled from most to least common.*](dep_sampling_diagram.png)
:::




##

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-cap: !expr glue("*ResStock 2025.1 AMY2018, upgrade 0, occupied units in Philadelphia. Outdoor weather sources from Philadelphia Airport 2018 & 20 year long term normals.*")
#| fig-width: 12
#| fig-height: 6.5
#| out-width: "100%"

# Legend kept on: the season ramp and the two outdoor line types need a key
philly_hi_plot +
  tsp_theme +
  theme(legend.position = "top")
```

## Takeaways

::: {.incremental}
1. These data are super cool, but they are simplifications designed specifically for energy consumption end points. Indoor environmental estimates are derived as a secondary consequence, and can be limited in some key ways.
2. Some information is better than no information, especially when it's derived in a tractable way that can be used to estimate uncertainty. 
3. Seems to me that we still need better observed data on housing and in the indoor environment. 
:::

## 

Thanks for listening!! 🙏